Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-20 arxiv_full_version/experiments.tex
arxiv version

\section{Simulation Studies}\label{sec_simulation}

@@

\textbf{Sample efficiency.}
First, we validate the result of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss as we increase the number of samples per-task   from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase distance $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

\textbf{Sample size ratio.}
Second, we validate the result of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$. 
The task-specific variance is $d^2 = 0.08$, the noise variance is $\sigma^2 = 4^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d^2 = 0.045$, task two's prediction loss decreases faster.
Intuitively, this is because bias increases less for smaller $d^2$.
%Our result from Corollary \ref{cor_MTL_loss} explains this trend.

\textbf{Covariate shift.}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.
%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

>>>>

\section{Numerical Results}

\subsection{Simulation Studies}\label{sec_simulation}

@@

\begin{figure*}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/same_covariates.eps}
		\caption{Example \ref{ex_same_cov}}
		\label{fig_same_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/sample_ratio_several_d.eps}
		\caption{Example \ref{ex_sample_ratio}}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/covariate_shift.eps}
		\caption{Example \ref{ex_covshift}}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	Our estimated losses (solid line) match the empirical losses (dotted line) accurately under various settings in dimension $p = 200$.
	\textbf{Left.} Validating Example \ref{ex_same_cov} for ten tasks: the noise variance $\sigma^2$ is $1/4$.
	\textbf{Middle.} Validating Example \ref{ex_sample_ratio} for two tasks: we discover an interesting phenomena by fixing task two's sample size and increasing task one's sample size.
	Moreover, our result accurately predicts the critical point (marked in circle) of the loss curve.
%	Depending on how large the distance $d^2$ is, task two's prediction loss decreases initially before increasing again, or decreases monotonically.
	\textbf{Right.} We show how different levels of covariate shift affect hard parameter sharing when there is no bias.
	Having covariate shift increases task two's prediction loss when task two's sample size is smaller than task one. Otherwise, having covariate shift (surprisingly) decreases task two's prediction loss.}
	\label{fig_model_shift_phasetrans}
\end{figure*}

\paragraph{Sample efficiency.}
First, we validate the result of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss over ten tasks as we increase the number of samples per-task from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend that the average prediction loss increases as we increase distance $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally, we find that HPS requires fewer samples per-task than STL to achieve the same loss level. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

\paragraph{Sample size ratio.}
Second, we validate the result of Example \ref{ex_sample_ratio}.
Figure \ref{fig_size} shows task two's prediction loss  as we increase the sample ratio $n_1 / n_2$ from $1/10$ to $7/10$.
%Again, our estimates are accurate compared to the empirical losses.
We consider a regime where task two consists of $80,000$ samples, and task one's sample size varies from $8,000$ to $56,000$. 
The task-specific variance (which scales with model distance) is $d = 0.2$, the noise variance is $\sigma^2 = 4^2$, and the shared signal variance is $1$. We observe that as we increase the sample ratio, task two's prediction loss decreases initially but later will increase when the sample ratio is above a certain level.
On the other hand, when $d = 0.15$, task two's prediction loss decreases faster.
Intuitively, this is because bias increases less for smaller $d^2$.
%Our result from Corollary \ref{cor_MTL_loss} explains this trend.

\paragraph{Covariate shift.}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.
%Covariate shift: We set $\kappa = 1$ and $d = 0$.
%We set $\rho_2 = 4$ and vary $\rho_1$ from $5$ to $25$ for sample sizes.
%We use the scale parameter $\lambda = 1$ for the curve without covariate shift and $\lambda = 2$ for the curve with covariate shift (cf. Section \ref{sec_covshift}).

### 2020-10-20 arxiv_full_version/intro.tex
arxiv version

% The idea is to use a shared feature space for all tasks, while each task also has a separate layer for making the prediction.
\begin{abstract}
	\vspace{-0.1in}
	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
	\vspace{-0.15in}
\end{abstract}

\vspace{-0.1in}
\section{Introduction}\label{sec_introduction}

@@

\begin{figure*}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.45\textwidth,valign=t]{figures/mtl_model_arch.pdf}
		\caption{A hard parameter sharing architecture}
		\label{fig_intro_arch}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\vspace{-0.1in}
	\caption{
	Left: an illustrative picture of HPS.
	Right: an illustrative example of using HPS for two tasks $X_1, Y_1$ and $X_2, Y_2$ with sample size $n_1, n_2$, respectively.
	Increasing $n_1/n_2$ decreases task two's prediction loss initially but increase afterward. This phenomenon occurs due to different bias-variance tradeoffs as $n_1/n_2$ increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
	\vspace{-0.07in}
\end{figure*}

>>>>

% The idea is to use a shared feature space for all tasks, while each task also has a separate layer for making the prediction.
\begin{abstract}
	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

\section{Introduction}\label{sec_introduction}

@@

\begin{figure*}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.45\textwidth,valign=t]{figures/mtl_model_arch.pdf}
		\caption{A hard parameter sharing architecture}
		\label{fig_intro_arch}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\vspace{-0.1in}
	\caption{
	Left: an illustrative picture of HPS.
	Right: an illustrative example of using HPS for two tasks $X_1, Y_1$ and $X_2, Y_2$ with sample size $n_1, n_2$, respectively.
	Increasing $n_1/n_2$ decreases task two's prediction loss initially but increase afterward. This phenomenon occurs due to different bias-variance tradeoffs as $n_1/n_2$ increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure*}

### 2020-10-20 arxiv_full_version/main.tex
arxiv version

%\usepackage[nonatbib]{neurips_2020}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\usepackage[export]{adjustbox}
\usepackage[margin=1in]{geometry}
%\usepackage{fullpage}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{macro_math}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage[us,12hr]{datetime}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{caption,subcaption}
%\usepackage{breqn}

% Squished list environment to save space
\newcommand{\squishlist}{
\begin{list}{$\bullet$}
{\setlength{\itemsep}{0pt}
\setlength{\parsep}{2pt}
\setlength{\topsep}{2pt}
\setlength{\partopsep}{0pt}
\setlength{\itemindent}{2pt}
\setlength{\leftmargin}{0.5cm}
% \setlength{\labelwidth}{1em}
}
}
\newcommand{\squishend}{\end{list}}

%\setlength{\parskip}{0.3em}
\begin{document}

%\title{Revisiting the Bias-Variance Tradeoff of Multi-Task Learning in High Dimensions}
\title{Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression}
\maketitle

>>>>

%\usepackage[nonatbib]{neurips_2020}

\def\shownotes{0}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\usepackage[export]{adjustbox}
\usepackage[margin=1in]{geometry}
%\usepackage{fullpage}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{macro_math}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage[us,12hr]{datetime}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{caption,subcaption}
%\usepackage{breqn}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

%\setlength{\parskip}{0.3em}

%\title{Revisiting the Bias-Variance Tradeoff of Multi-Task Learning in High Dimensions}
\title{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}
\usepackage{authblk}
\author[$\dagger$]{Hongyang R. Zhang\thanks{Equal contribution. Correspondence to hrzhang@northeastern.edu.}}
\author[$\ddagger$]{Fan Yang\samethanks}
\author[$\mathsection$]{Sen Wu}
\author[$\ddagger$]{Weijie S. Su}
\author[$\mathsection$]{Christopher R\'e}
\affil[$\dagger$]{Khoury College of Computer Sciences, Northeastern University}
\affil[$\ddagger$]{Department of Statistics, Wharton School, University of Pennsylvania}
\affil[$\mathsection$]{Department of Computer Science, Stanford University}

\begin{document}
\maketitle
